---
title: "Untitled"
format: html
editor: visual
---

Extract data for twensday


Tyler L. McIntosh
CU Boulder CIRES Earth Lab
Last updated: 1/24/24

This script uses the following naming conventions wherever possible:
 lowerCamelCase for variables
 period.separated for functions
 underscore_separated for files

cp -r ~/data-store/data/iplant/home/shared/earthlab/twensday3_transitionwarningsigns /home/jovyan/twensday3_transitionwarningsigns

cp -r /home/jovyan/twensday3_transitionwarningsigns/data/derived ~/data-store/data/iplant/home/shared/earthlab/twensday3_transitionwarningsigns/data/derived

# Setup workspace
User-set global parameters
Directory management
Package management

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}

rm(list=ls()) #Ensure empty workspace if running from beginning

#################################################
#####EVERYTHING HERE SHOULD BE SET MANUALLY######
#################################################

computing <- "cyverse" #Sets computing location and file setup; "cyverse" or "local"
nCores <- 15 #Number of cores available - only necessary if on CyVerse (future package struggles to read automatically on CyVerse, reads 128 (the maximum number of cores)


#################################################

# Directory management ----


#Set directories
if(computing == "local") {
  home <- "C:/Users/tyler/OneDrive - UCB-O365/dev/state-transitions-early-warning"
} else if(computing == "cyverse") {
  home <- "./twensday3_transitionwarningsigns"
}


# Package management ----

#source(here::here('code/functions.R'), local = FALSE)

#Check the required libraries and download if needed
packageList <- c(
  "tidyverse", #Includes ggplot2, dplyr, tidyr, readr, purrr, tibble, stringr, forcats
  "terra", #New raster data package, documentation pdf here: https://cran.r-project.org/web/packages/terra/terra.pdf
  "future", "future.apply", "furrr", "doFuture", "progressr", #Futureverse! https://www.futureverse.org/packages-overview.html; https://henrikbengtsson.github.io/course-stanford-futureverse-2023/
  "parallelly", #some useful functions from 'future' were moved to 'parallelly' because they were helpful outside the future framework
  "sf", #New vector data package
  #"mapview", #For quick interactive mapping
  "tictoc", #time running of processes
  "glue", #easy strings
  "tigris", #state polygons
  "data.table",
  "vroom",
  "ggpmisc"  #For adding R^2 to plots
)


# A function to check and install packages provided to the function. Written in part by ChatGPT4.
# PARAMETERS
# packageList : a vector of packages used in the script, e.g. c("here", "dplyr")
# autoInstall : either "y" or "n" (default). "y" will install all required packages without asking the user for permission. "n" will ask permission from the user.
install.and.load.packages <- function(packageList, autoInstall = "n") {
  missingCranPackages <- c()
  missingGithubPackages <- c()
  for (package in packageList) {
    packageLoaded <- FALSE
    # Check if the package is from GitHub
    if (grepl("/", package)) {
      packageName <- unlist(strsplit(package, "/"))[2]
      packageLoaded <- require(packageName, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
    } else {
      packageLoaded <- require(package, character.only = TRUE, quietly = TRUE, warn.conflicts = FALSE)
    }
    # Mark as missing if not loaded
    if (!packageLoaded) {
      if (grepl("/", package)) {
        missingGithubPackages <- c(missingGithubPackages, package)
      } else {
        missingCranPackages <- c(missingCranPackages, package)
      }
    }
  }
  # Install missing CRAN packages
  if (length(missingCranPackages) > 0) {
    cat("The following CRAN packages are missing: ", paste(missingCranPackages, collapse = ", "), "\n")
    if(autoInstall == "y") {
      response <- "y"
      cat("Installing the missing packages\n")
    } else {
      response <- readline(prompt = "\nDo you want to install the missing CRAN packages? (y/n): ")
    }
    if (tolower(response) == "y") {
      install.packages(missingCranPackages)
      for (package in missingCranPackages) {
        require(package, character.only = TRUE)
      }
    } else {
      cat("Skipping installation of missing CRAN packages.\n")
    }
  }
  # Ask to install the 'remotes' package if GitHub packages are missing and 'remotes' is not installed
  if (length(missingGithubPackages) > 0 && !requireNamespace("remotes", quietly = TRUE)) {
    if(autoInstall == "y") {
      response <- "y"
      cat("Installing 'remotes' package to install GitHub packages\n")
    } else {
      response <- readline(prompt = "\nDo you want to install the 'remotes' package? (It is required to install packages from GitHub) (y/n): ")
    }
    if (tolower(response) == "y") {
      install.packages("remotes")
    } else {
      cat("Skipping installation of GitHub packages.\n")
      missingGithubPackages <- c() # Clear the list of GitHub packages
    }
  }
  # Install missing GitHub packages
  if (length(missingGithubPackages) > 0) {
    cat("The following GitHub packages are missing: ", paste(missingGithubPackages, collapse = ", "), "\n")
    if(autoInstall == "y") {
      response <- "y"
      cat("Installing the missing packages\n")
    } else {
      response <- readline(prompt = "\nDo you want to install the missing GitHub packages? (y/n): ")
    }    
    if (tolower(response) == "y") {
      for (package in missingGithubPackages) {
        remotes::install_github(package)
        packageName <- unlist(strsplit(package, "/"))[2]
        require(packageName, character.only = TRUE)
      }
    } else {
      cat("Skipping installation of missing GitHub packages.\n")
    }
  }
  cat("All specified packages installed and loaded.\n")
}

install.and.load.packages(packageList, autoInstall = "y")



#Start parallel computing ----
#Set cores
# if(computing == "local") {
#   nCores <- future::availableCores() - 1
# } else if(computing == "cyverse") {
#   cat("Using user-defined number of computing cores:", nCores)
# }
# future::plan("multisession", workers = nCores)

```


# Load data paths

```{r load-data, echo = FALSE}



#Specific derived data directory for this script's outputs
outDir <- file.path(home, "data/derived")
if (!dir.exists(outDir)){
  dir.create(outDir)
}


#Load data

#Raw data
dStackF <- list.files(file.path(home, 'data/raw/western-conus'), pattern = "forest-disturbance-stack", full.names = TRUE)
srConversionsF <- file.path(home, "data/shared/SR_conversion/transformations_data_from_1999_2021.csv")



dStack <- terra::rast(dStackF)

#Function to create disturbance layer names
re.name.stack <- function(dStack, dNamesRoot) {
  dNums <- as.character(seq(1999, 2020))
  dNames <- paste(dNamesRoot, dNums, sep="_")
  names(dStack) <- dNames
  return(dStack)
}

dStack <- dStack |> re.name.stack("forestDisturbance")


```


# Prep data and output for later analysis
```{r}

nms <- names(readr::read_csv(srConversionsF, n_max = 2))

oneCol <- vroom::vroom(srConversionsF, col_select = 1)
nrows <- nrow(oneCol)
rm(oneCol)
gc()

nrows / 500000

fireCodes <- c(1,5,9,13)
insectCodes <- c(2,6,10,14)
otherCodes <- c(3,7,11,15)
hotterDOnlyCodes4Plus <- c(4,8,12)
comboCodes <- c(5,6,7,9,10,11,13,14,15)
anyHotterDCodes <- seq(4,15)


##FUNCTION TO DO OPERATION
pull.disturbance.data.at.transitions <- function(skip, nmax, dStack) {

  if(class(dStack)[1] == "PackedSpatRaster") {
    dStack <- terra::unwrap(dStack)
  }

  workingSet <- vroom::vroom(srConversionsF, skip = skip, n_max = nmax, col_names = FALSE)|>
    `names<-`(nms) 
  
  points = sf::st_as_sf(workingSet, coords = c("easting", "northing"), crs = 5070) |> sf::st_transform(terra::crs(dStack))
  #mapview::mapview(points)
  
  extracted <- terra::extract(dStack, points)

  rm(dStack)
  

  fullDats <- cbind(workingSet, extracted) |>
    dplyr::mutate(transformType = dplyr::case_when(trees_to_grass > 0 & grass_to_trees > 0 & transformed == 0 ~ "shortTransform",
                                                   trees_to_grass > 12 & grass_to_trees == 0 ~ "transformNeverBackShort",
                                                   trees_to_grass > 0 & trees_to_grass <= 12 & grass_to_trees == 0 ~ "transformNeverBack10Plus",
                                                   transformed == 1 ~ "transformBackAfter10Plus",
                                                   TRUE ~ "noTransform"),
                  firstYearTransformed = 2000 + trees_to_grass)
  df <- fullDats

  out <- df %>%
    mutate(
      disturbance1YearPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 1)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)  # In case the year_col_name does not exist
        }
      }),
      disturbance2YearsPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 2)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)
        }
      }),
      disturbance3YearsPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 3)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)
        }
      }),
      disturbance4YearsPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 4)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)
        }
      }),
      disturbance5YearsPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 5)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)
        }
      }),
      transformedTooEarly = if_else(trees_to_grass > 0 & trees_to_grass < 4, TRUE, FALSE),
      doNotProcess = if_else(transformedTooEarly | transformType == "noTransform", TRUE, FALSE),
      yearPriorIncident = dplyr::case_when(is.na(disturbance1YearPrior) ~ NA,
                                           disturbance1YearPrior %in% fireCodes ~ "fire",
                                           disturbance1YearPrior %in% insectCodes ~ "insect",
                                           disturbance1YearPrior %in% otherCodes ~ "other",
                                           disturbance1YearPrior %in% hotterDOnlyCodes4Plus ~ "hotterDrought",
                                           disturbance1YearPrior %in% c(0) ~ "none"),
      
      
      anyHotDInYearsPrior = case_when(doNotProcess ~ NA,
                                      TRUE ~ pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% anyHotterDCodes))),
      onlyHotDInYearsPrior = case_when(doNotProcess ~ NA,
                                      TRUE ~ pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% hotterDOnlyCodes4Plus))),
      anyFireInYearsPrior = case_when(doNotProcess ~ NA,
                                      TRUE ~ pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% fireCodes))),
      anyInsectInYearsPrior = case_when(doNotProcess ~ NA,
                                      TRUE ~ pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% insectCodes))),
      combosInYearsPrior = case_when(doNotProcess ~ NA,
                                      TRUE ~ pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% comboCodes)))

      
      
    #   anyHotDInYearsPrior = if_else(doNotProcess, NA_integer_, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% anyHotterDCodes))),
    #   onlyHotDInYearsPrior = if_else(doNotProcess, NA_integer_, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% hotterDOnlyCodes4Plus))),
    #   anyFireInYearsPrior = if_else(doNotProcess, NA_integer_, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% fireCodes))),
    #   anyInsectInYearsPrior = if_else(doNotProcess, NA_integer_, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% insectCodes))),
    #    combosInYearsPrior = if_else(doNotProcess, NA_integer_, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% comboCodes)))#, NOTE: THESE MUTATE COMMANDS REMOVED SINCE THEY ARE OPERATING ROWWISE AND ARE INEFFICIENT. A VECTORIZED OPERATIONS ALTERNATIVE IS USED INSTEAD BETWEEN HERE AND THE END OF THE FUNCTION
    #   # totalHotDInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% anyHotterDCodes)),
    #   # totalOnlyHotDInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% hotterDOnlyCodes4Plus)),
    #   # totalFireInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% fireCodes)),
    #   # totalInsectInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% insectCodes)),
    #   # totalCombosInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% comboCodes))
    
    )

selected_columns <- df %>%
  select(starts_with("forestDisturbance_"))

# Calculate the totals by vectorized operations
out$totalHotDInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% anyHotterDCodes), na.rm = TRUE)
out$totalOnlyHotDInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% hotterDOnlyCodes4Plus), na.rm = TRUE)
out$totalFireInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% fireCodes), na.rm = TRUE)
out$totalInsectInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% insectCodes), na.rm = TRUE)
out$totalCombosInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% comboCodes), na.rm = TRUE)

    
  vroom::vroom_write(x = out, file = file.path(outDir, paste0("transform_with_disturbance_", skip, ".csv")), delim = ",")
}


# tic()
# pull.disturbance.data.at.transitions(skip = 1, nmax = 500000, dStack = dStack)
# toc()


tic()
chunkSize = 1000000
nChunks <- nrows %/% chunkSize + 1
skips <- seq(from = 1, by = chunkSize, length.out = nChunks)

future::plan("multisession", workers = nCores)
options(future.globals.onReference = "error")

wrappedDStack <- terra::wrap(dStack)
skips[1:2] |> furrr::future_map(.f = ~pull.disturbance.data.at.transitions(skip = .x, nmax = chunkSize, dStack = wrappedDStack))
toc()



pull.disturbance.data.at.transitions(skip = skips[1], nmax = chunkSize, dStack = wrappedDStack)


```


# Analyze outputs

```{r}

outFs <- list.files(outDir, full.names = TRUE)
ttt <- vroom(outFs[1])

allDats <- vroom::vroom(outFs)

allDats <- vroom::vroom(outFs[3:96])

view <- allDats[1:50,]


allDats <- allDats |>
      dplyr::mutate(
        transformedTooEarly = if_else(trees_to_grass > 0 & trees_to_grass < 4, TRUE, FALSE),
        doNotProcess = if_else(transformedTooEarly | transformType == "noTransform", TRUE, FALSE),
        yearPriorIncident = dplyr::case_when(is.na(disturbance1YearPrior) ~ NA,
                                             disturbance1YearPrior %in% fireCodes ~ "fire",
                                             disturbance1YearPrior %in% insectCodes ~ "insect",
                                             disturbance1YearPrior %in% otherCodes ~ "other",
                                             disturbance1YearPrior %in% hotterDOnlyCodes4Plus ~ "hotterDrought",
                                             disturbance1YearPrior %in% c(0) ~ "none")
      )

removeDNP <- allDats |> dplyr::filter(!doNotProcess)

t <- removeDNP |> filter(trans_back != 0)
hist(t$trans_back)



ggplot2::ggplot() +
  ggplot2::geom_histogram(data = removeDNP, mapping = aes(x = firstYearTransformed), binwidth = 1) +
  theme_minimal() +
  labs(x = "First year transformed", y = "Frequency", title = "Year of transformation")

ggplot2::ggplot() +
  ggplot2::geom_histogram(data = removeDNP, mapping = aes(x = firstYearTransformed), binwidth = 1) +
  facet_wrap(~ transformType) +
  theme_minimal() +
  labs(x = "First year transformed", y = "Frequency", title = "Year of transformation by transformation type")



ggplot2::ggplot() +
  ggplot2::geom_bar(data = removeDNP, mapping = aes(x = yearPriorIncident), stat="count", position = "dodge") +
  theme_minimal() +
  labs(x = "Disturbance type", y = "Count", title = "Disturbance in year prior to transformation")

ggplot2::ggplot() +
  ggplot2::geom_bar(data = removeDNP, mapping = aes(x = yearPriorIncident), stat="count", position = "dodge") +
  facet_wrap(~ transformType) +
  theme_minimal() +
  labs(x = "Disturbance type", y = "Count", title = "Disturbance in year prior to transformation by transformation type")


nonFireTransition <- removeDNP |> dplyr::filter(yearPriorIncident == "none")


moreDats <- nonFireTransition %>%
  dplyr::mutate(anyHotDInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% anyHotterDCodes))),
      onlyHotDInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% hotterDOnlyCodes4Plus))),
      anyFireInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% fireCodes))),
      anyInsectInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% insectCodes))),
      combosInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% comboCodes))))


ggplot2::ggplot() +
  ggplot2::geom_histogram(data = moreDats, mapping = aes(x = onlyHotDInYearsPrior ), binwidth = 1) +
  theme_minimal() +
  labs(x = "Hotter drought only in 5 years prior", y = "Frequency", title = "Hotter drought prior to transformation when \n no disturbance in year prior")

ggplot2::ggplot() +
  ggplot2::geom_histogram(data = moreDats, mapping = aes(x = onlyHotDInYearsPrior ), binwidth = 1) +
  facet_wrap(~ transformType) +
  theme_minimal() +
  labs(x = "Hotter drought only in 5 years prior", y = "Frequency", title = "Hotter drought prior to transformation when \n no disturbance in year prior, by transform type")

ggplot2::ggplot() +
  ggplot2::geom_histogram(data = moreDats, mapping = aes(x = anyFireInYearsPrior), binwidth = 1) +
  theme_minimal() +
  labs(x = "Fire in 5 years prior", y = "Frequency", title = "Fire prior to transformation when no disturbance in year prior")

ggplot2::ggplot() +
  ggplot2::geom_histogram(data = moreDats, mapping = aes(x = anyInsectInYearsPrior), binwidth = 1) +
  theme_minimal() +
  labs(x = "Insects in 5 years prior", y = "Frequency", title = "Insects prior to transformation when no disturbance in year prior")





tic()
df <- tSet
df <- df[1:50000,]
df <- df %>%
    mutate(
      disturbance1YearPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 1)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)  # In case the year_col_name does not exist
        }
      }),
      disturbance2YearsPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 2)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)
        }
      }),
      disturbance3YearsPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 3)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)
        }
      }),
      disturbance4YearsPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 4)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)
        }
      }),
      disturbance5YearsPrior = sapply(1:n(), function(i) {
        year_col_name <- paste0("forestDisturbance_", df$firstYearTransformed[i] - 5)
        if (year_col_name %in% names(df)) {
          value <- as.integer(df[[year_col_name]][i])
          return(value)
        } else {
          return(NA_integer_)
        }
      }),
      transformedTooEarly = if_else(trees_to_grass > 0 & trees_to_grass < 4, TRUE, FALSE),
      doNotProcess = if_else(transformedTooEarly | transformType == "noTransform", TRUE, FALSE),
      yearPriorIncident = dplyr::case_when(is.na(disturbance1YearPrior) ~ NA,
                                           disturbance1YearPrior %in% fireCodes ~ "fire",
                                           disturbance1YearPrior %in% insectCodes ~ "insect",
                                           disturbance1YearPrior %in% otherCodes ~ "other",
                                           disturbance1YearPrior %in% hotterDOnlyCodes4Plus ~ "hotterDrought",
                                           disturbance1YearPrior %in% c(0) ~ "none"),
      anyHotDInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% anyHotterDCodes))),
      onlyHotDInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% hotterDOnlyCodes4Plus))),
      anyFireInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% fireCodes))),
      anyInsectInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% insectCodes))),
      combosInYearsPrior = if_else(doNotProcess, NA, pmap_int(select(., disturbance1YearPrior:disturbance5YearsPrior), ~ sum(c(...) %in% comboCodes)))#,
      # totalHotDInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% anyHotterDCodes)),
      # totalOnlyHotDInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% hotterDOnlyCodes4Plus)),
      # totalFireInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% fireCodes)),
      # totalInsectInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% insectCodes)),
      # totalCombosInAllYears = pmap_int(select(., forestDisturbance_1999:forestDisturbance_2020), ~ sum(c(...) %in% comboCodes))
    )

selected_columns <- df %>%
  select(starts_with("forestDisturbance_"))

# Calculate the totals by vectorized operations
df$totalHotDInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% anyHotterDCodes), na.rm = TRUE)
df$totalOnlyHotDInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% hotterDOnlyCodes4Plus), na.rm = TRUE)
df$totalFireInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% fireCodes), na.rm = TRUE)
df$totalInsectInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% insectCodes), na.rm = TRUE)
df$totalCombosInAllYears <- rowSums(sapply(selected_columns, function(x) x %in% comboCodes), na.rm = TRUE)

toc()
